---
title: "Database data science"
output: html_notebook
---

```{r setup}
# Packages ----
library(dplyr)
library(dbplyr)
library(DBI)
library(dbplot)
library(ggplot2)
library(tidypredict)
library(modeldb)
library(nycflights13)
library(parsnip)
```

# Setup
Establish a connection to the database. For this example, an in memory SQLite
database created through the `RSQLite` package.
```{r connection}
con <- dbConnect(RSQLite::SQLite(), ":memory:")
```

Transfer some data into the database.
```{r}
dbWriteTable(con, "flights", flights)
dbWriteTable(con, "airports", airports)
dbWriteTable(con, "airlines", airlines)
dbWriteTable(con, "planes", planes)
dbWriteTable(con, "weather", weather)
dbWriteTable(con, "cars", mtcars)
```

# Import, Tidy, and Transform
Create reference to a database table.
```{r}
flights_db <- tbl(con, "flights")
class(flights_db)
```

What happens when I run take the first 6 rows of `flights_db`?
```{r}
flights_db %>% 
  head()
```

```{r}
flights_db %>% 
  head() %>% 
  show_query()
```

See how the generated SQL code changes from one provider to another.
```{r}
flights_db %>% 
  head() %>% 
  sql_render(con = simulate_mssql())
```

Calculate average delays by carrier.
```{r}
delays <- flights_db %>% 
  group_by(carrier) %>% 
  summarise(total_flights = n(),
            avg_dep_delay = mean(dep_delay, na.rm = TRUE),
            avg_arr_delay = mean(arr_delay, na.rm = TRUE)) %>% 
  mutate(avg_delay_diff = avg_arr_delay - avg_dep_delay) %>% 
  arrange(desc(avg_delay_diff))
```

```{r}
delays %>% 
  show_query()
```

```{r}
class(delays)
```

```{r}
results <- collect(delays)
class(results)
results
```

Join carrier details
```{r}
airlines_db <- tbl(con, "airlines")
```

```{r}
detailed_delays <- delays %>% 
  left_join(airlines_db) %>% 
  select(name, everything())
```

```{r}
detailed_delays %>% 
  show_query()
```

```{r}
detailed_delays
```

# Visualize
Create bar plot of carrier counts.
```{r}
p <- flights_db %>% 
  dbplot_bar(carrier)

p
```

```{r}
p +
  theme_minimal() +
  labs(title = "Carrier Counts",
       y = "Count")
```

```{r}
flights_db %>% 
  db_compute_count(carrier)
```

# Model
Define a model in R.
```{r}
lm_fit <- lm(mpg ~ hp + cyl, data = mtcars)
lm_fit
```

Translate the R model into SQL using the `tidypredict` package.
```{r}
tidypredict_sql(lm_fit, con)
```

Predict in the database.
```{r}
cars_db <- tbl(con, "cars")
tidypredict_to_column(cars_db, lm_fit) %>% 
  select(mpg, fit)
```

Fit a model in the database using the `modeldb` package.
```{r}
db_fit <- cars_db %>% 
  select(mpg, hp, cyl) %>% 
  linear_regression_db(mpg)

db_fit
```
